---
title: "Response bias modelling"
format: html
editor: visual
html:
    toc: true
    toc-location: right
    toc-expand: true
    code-fold: true
    mathjax: true

project:
  type: website
  output-dir: docs

execute:
  freeze: auto
---

## Mixture modelling on core trials

*Core* trials made up the majority (90%; 432/480 trials) of the task. They're what we've discussed in meetings so far: participants saw an equal number of trials from each cell of our 2 x 2 design (2 reward levels: low, high; 2 coherence levels: 20%, 40%).

Here, we do a mixture model: we assume that for some percentage of trials, each responder is making a genuine effort at the task, but for others, they are guessing at random. This modelling approach estimates the proportion of trials which fall into each of these categories, then estimates our parameters of interest (value bias, precision) only for the distribution of 'genuine effort' trials.

On this run I left in the people who we usually exclude as frequent guessers, to see if the model still fits well (by assigning them a high guessing probability).

### Does the model fit?

First, let's visually inspect how the model is fitting relative to the observed data.

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing

pacman::p_load(brms, emmeans, bayestestR,	tidybayes, dplyr, ggplot2, bayesplot, ggh4x, knitr, kableExtra)

df=read.csv("exptrialdata_merged.csv")

df$coherence=as.factor(df$coherence)
df$trialval=as.factor(df$trialval)

df$valuebias_rad = (df$valuebias*pi/180)
df$colourangle_rad = (df$colourangle*pi/180)
df$greyangle_rad = (df$greyangle*pi/180)
df$response_rad = (df$response*pi/180)
df$id_orig = df$id
df$id = as.numeric(factor(df$pnum))

df$block = as.factor(ceiling(df$trialnum/60))

core = filter(df, trialtype=="core" & soft_timeout_flag == 0)
combo = filter(df, trialtype!="comp" & soft_timeout_flag == 0)

##### Filtering participants #####

trial_exclusions <- core %>%
  count(id) %>%
  mutate(droprate = 1-(n / 432))

exclusion_drops <- trial_exclusions %>%
  filter(droprate > 0.2) %>%
  pull(id)

rayleigh_results <- core %>%
  group_by(id) %>%
  summarise(
    n = n(),
    rayleigh_p = tryCatch({
      rayleigh.test(circular(valuebias_rad))$p.value
    }, error = function(e) NA),
    rayleigh_stat = tryCatch({
      rayleigh.test(circular(valuebias_rad))$statistic
    }, error = function(e) NA),
    .groups = "drop"
  )

rayleigh_drops <- rayleigh_results %>%
  filter(rayleigh_p > 0.05) %>%
  pull(id)

drops <- union(exclusion_drops, rayleigh_drops)

combo_filt <- combo %>%
  filter(!(id %in% drops)) %>%
  mutate(
    trialval = droplevels(as.factor(trialval)),
    coherence = droplevels(as.factor(coherence))
    )

# 3. Generate y_rep

core_filt <- core %>%
  filter(!(id %in% c(26))) %>%
  mutate(
    trialval = droplevels(as.factor(trialval)),
    coherence = droplevels(as.factor(coherence))
    )

fit_combo_dist <- readRDS("fit_combo_dist_3x2.rds")
fit_mixture_test <- readRDS("dotmotion_mixmodel.rds")

valid_ids <- rownames(ranef(fit_combo_dist)$id)

# 2. Filter the data to match the model's knowledge
# This ensures y_obs and y_rep are the same length and refer to the same people
combo_filt_clean <- combo_filt %>%
  mutate(id = as.character(id)) %>%
  filter(id %in% valid_ids) %>%
  mutate(
    trialval = droplevels(as.factor(trialval)),
    coherence = droplevels(as.factor(coherence)),
    condition = interaction(trialval, coherence, sep = " | Coh: ")
  )


y_rep_mix <- posterior_predict(fit_mixture_test, ndraws = 10)
y_obs <- core_filt$valuebias_rad

# 2. Create the interaction for paneling
core_filt$condition <- interaction(core_filt$trialval, 
                                   core_filt$coherence, 
                                   sep = " | Coh: ")

# 3. Grouped PPC plot
library(bayesplot)

ppc_dens_overlay_grouped(
  y = y_obs * 180/pi, 
  yrep = y_rep_mix * 180/pi, 
  group = core_filt$condition
) + 
  theme_minimal() +
  labs(
    title = "Mixture Model PPC by Condition",
    subtitle = "Dark line = Data | Light lines = Mixture of Von Mises + Uniform simulations",
    x = "Value bias (degrees)",
    y = "Density"
  ) +
  facet_wrap(~ group, ncol = 2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", alpha = 0.5)

```

Overall, model fit is looking much better than for the simpler non-mixture version (circular response distribution with dispersion.

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing

y_rep <- posterior_predict(
  fit_combo_dist, 
  newdata = combo_filt_clean,
  ndraws = 10
)

# 4. Get the observed values from the clean dataset
y_obs <- combo_filt_clean$valuebias_rad

# 5. Plot
ppc_dens_overlay_grouped(
  y = y_obs * 180 / pi, 
  yrep = y_rep * 180 / pi, 
  group = combo_filt_clean$condition
) + 
  theme_minimal() +
  labs(
    title = "Posterior Predictive Check by Condition",
    x = "Value bias (°)",
    y = "Density"
  ) +
  facet_wrap(~ group, ncol = 2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black", alpha = 0.5)

```

### Is a response bias present?

First let's see how the model has estimated *mu* (mean) in each reward x coherence condition.

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing

cell_draws_mix <- emmeans(fit_mixture_test, ~ trialval * coherence, dpar = "mu1") %>%
  gather_emmeans_draws()

# 2. Test vs. 0: Is the task-directed bias credible?
zero_test_summary_mix <- cell_draws_mix %>%
  group_by(trialval, coherence) %>%
  summarize(
    # Calculations in radians first
    median_rad = median(.value),
    lower_rad  = quantile(.value, 0.025),
    upper_rad  = quantile(.value, 0.975),
    # p_direction returns the probability that the effect is strictly >0 or <0
    pd = as.numeric(p_direction(.value)),
    .groups = "drop"
  ) %>%
  # Convert to degrees for human readability
  mutate(
    median_deg = round(median_rad * 180 / pi, 2),
    lower_95_deg = round(lower_rad * 180 / pi, 2),
    upper_95_deg = round(upper_rad * 180 / pi, 2),
    pd = round(pd, 3)
  )

# 3. Create the Kable Table

# 4. Visualization
ggplot(zero_test_summary_mix, aes(x = trialval, y = median_deg, color = trialval)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", alpha = 0.6) +
  geom_pointrange(aes(ymin = lower_95_deg, ymax = upper_95_deg), 
                  size = 1, linewidth = 1.2) +
  facet_wrap(~ coherence, labeller = label_both) +
  scale_color_manual(values = c("hi" = "#ff6117", "lo" = "#009dff")) +
  labs(
    title = "Value bias (mu) for non-guess responses",
    x = "Trial reward",
    y = "Value bias (°)",
    color = "Reward Level"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.minor = element_blank(),
        strip.text = element_text(face = "bold"),
        legend.position = 'none')

zero_test_summary_mix %>%
  select(trialval, coherence, median_deg, lower_95_deg, upper_95_deg, pd) %>%
  knitr::kable(
    caption = "Value bisa yb condition",
    col.names = c("Reward", "Coherence", "Median Bias", "95% Lower", "95% Upper", "P-Directional"),
    align = "lccccc"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

Across all value x coherence combinations, there is a credible response bias. However, the 'odd one out' condition is low coherence/high reward, where the bias is substantially larger than for other conditions.

### How does bias change with coherence and reward?

Here, we check the posterior distributions for *differences* in response bias across levels of coherence and reward (i.e., the range of plausible values for how much the response has actually shifted). In these graphs, the plotted point is the median difference and the whiskers are the 66% (thick) and 95% (thin) credible intervals.

#### Effect of coherence

Firstly, let's take a look at how coherence affects response bias within each level of reward.

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing

all_draws_mix <- emmeans(fit_mixture_test, pairwise ~ trialval * coherence, dpar = "mu1")$contrasts %>%
  gather_emmeans_draws()

# 2. Define the relevant coherence contrasts (0.2 vs 0.4)
# Note: emmeans usually formats these as "hi 0.2 - hi 0.4"
ordered_levels_mix <- c(
  "hi 0.2 - hi 0.4",
  "lo 0.2 - lo 0.4"
)

coherence_effects_mix <- all_draws_mix %>%
  filter(
    # Matches "hi ... - hi ..."
    (grepl("hi", contrast) & !grepl("lo", contrast)) | 
      # Matches "lo ... - lo ..."
      (grepl("lo", contrast) & !grepl("hi", contrast))
  ) %>%
  # Now that we only have within-reward rows, check the coherence
  filter(grepl("0.2", contrast) & grepl("0.4", contrast)) %>%
  mutate(
    reward_group = ifelse(grepl("hi", contrast), "High Reward", "Low Reward"),
    diff_deg = .value * 180 / pi
  )

# 2. This should now be 4000 and 4000 (or whatever your ndraws is)
# print(table(coherence_effects_mix$reward_group))

# 3. Plot
ggplot(coherence_effects_mix, aes(x = diff_deg, y = reward_group, fill = reward_group)) +
  stat_halfeye(alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_fill_manual(values = c("High Reward" = "#ff6117", "Low Reward" = "#009dff")) +
  labs(
    title = "Effect of coherence upon value bias (within reward levels)",
    x = "Value bias difference (low - high coherence)",
    y = ""
  ) +
  theme_minimal()

# 5. Summary Table
coherence_diff_summary_mix <- coherence_effects_mix %>%
  group_by(contrast) %>%
  summarize(
    median = median(diff_deg),
    lower_95 = quantile(diff_deg, 0.025),
    upper_95 = quantile(diff_deg, 0.975),
    pd = as.numeric(bayestestR::p_direction(.value)),
    .groups = "drop"
  ) %>%
  arrange(match(contrast, ordered_levels_mix)) 

coherence_diff_summary_mix %>%
  mutate(across(where(is.numeric), ~ round(., 3))) %>%
  knitr::kable(
    format = "html",
    caption = "Change in value bias across coherence levels",
    col.names = c("Condition Contrast", "Median Diff", "Lower 95%", "Upper 95%", "P-Direction"),
    align = "lcccc"
  ) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

Value bias decreases with increasing coherence in the high-reward condition only; there's no meaningful difference across coherence levels for low-reward trials, as in neither case is there that much of a bias to start with.

#### Effect of reward

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing

draws_clean <- emmeans(fit_mixture_test, pairwise ~ trialval * coherence, dpar = "mu1")$contrasts %>%
  gather_emmeans_draws() %>%
  mutate(diff_deg = .value * 180 / pi)

# 2. Filter specifically for Reward Effects (Hi - Lo) within Coherence levels
reward_effect_clean <- draws_clean %>%
  filter(
    # Rule A: Must compare Hi to Lo
    (grepl("hi", contrast) & grepl("lo", contrast)) & 
      # Rule B: Must happen within 0.2 OR within 0.4
      (grepl("0.2", contrast) | grepl("0.4", contrast)) &
      # Rule C: BAN mixing coherence levels (0.2 vs 0.4)
      !(grepl("0.2", contrast) & grepl("0.4", contrast))
  ) %>%
  mutate(
    # Create nice labels for the plot
    coherence_level = ifelse(grepl("0.2", contrast), "Low Coherence (0.2)", "High Coherence (0.4)"),
    coherence_level = factor(coherence_level, levels = c("High Coherence (0.4)", "Low Coherence (0.2)"))
  )

# 3. Plot with the consistent Y-axis
ggplot(reward_effect_clean, aes(x = diff_deg, y = coherence_level, fill = coherence_level)) +
  stat_halfeye(alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_fill_manual(values = c("Low Coherence (0.2)" = "#807dba", "High Coherence (0.4)" = "#bcbddc")) +
  labs(
    title = "Effect of reward upon value bias (within coherence levels)",
    x = "Value bias difference (high - low reward)",
    y = ""
  ) +
  theme_minimal(base_size = 14) +
  guides(fill = "none")

# Summary Table for the Task-Directed Bias Shift (mu1)
reward_diff_summary_mix <- reward_effect_clean %>%
  group_by(coherence_level) %>%
  summarize(
    median = median(diff_deg),
    lower_95 = quantile(diff_deg, 0.025),
    upper_95 = quantile(diff_deg, 0.975),
    # p_direction (pd) represents the certainty of the effect direction
    pd = as.numeric(bayestestR::p_direction(.value)),
    .groups = "drop"
  ) %>%
  # Ensures 0.2 (Low Coherence) stays at the top of the table
  arrange(desc(coherence_level))

# Render the HTML Table
reward_diff_summary_mix %>%
  mutate(across(where(is.numeric), ~ round(., 2))) %>%
  knitr::kable(
    format = "html",
    caption = "Change in value bias across reward levels (high - low)",
    col.names = c("Coherence Level", "Median Diff", "95% Lower", "95% Upper", "P-Direction"),
    align = "lcccc"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE,
    position = "left"
  )
```

Value bias is greater in the high-reward condition than the low-reward one, only when coherence is low.

### Does response precision change across trial types?

Next, let's see how the model has estimated *kappa* (response precision) in each reward x coherence condition.

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing

ordered_coherence <- c("0.2", "0.4") 

# 2. Extract draws for the task-directed precision parameter (kappa1)
precision_estimates_mix <- emmeans(fit_mixture_test, ~ trialval * coherence, dpar = "kappa1") %>%
  gather_emmeans_draws() %>%
  mutate(
    # emmeans/brms returns kappa on the log scale for mixture models
    actual_kappa = exp(.value), 
    coherence = factor(coherence, levels = ordered_coherence),
    trialval = factor(trialval, levels = c("hi", "lo"))
  )

# 3. The Mirrored Plot
ggplot(precision_estimates_mix, aes(x = actual_kappa, y = coherence, fill = trialval)) +
  stat_halfeye(
    # Mirrored layout: Hi on top, Lo on bottom
    aes(side = ifelse(trialval == "hi", "top", "bottom")),
    scale = 0.7,
    alpha = 0.8,
    point_interval = median_qi,
    position = position_dodge(width = 0.3) 
  ) +
  scale_fill_manual(
    values = c("hi" = "#ff6117", "lo" = "#009dff"),
    labels = c("hi" = "High reward", "lo" = "Low reward")
  ) +
  scale_y_discrete(labels = c(
    "0.2" = "Low (0.2)", 
    "0.4" = "High (0.4)"
  )) +
  labs(
    x = "Concentration parameter (kappa) for non-guess responses",
    y = "Coherence",
    fill = "Reward"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_line(color = "grey95"),
    axis.text.y = element_text(face = "bold")
  )

# 4. Summary Table
precision_summary_mix <- precision_estimates_mix %>%
  group_by(coherence, trialval) %>%
  summarize(
    median_kappa = median(actual_kappa),
    lower_95 = quantile(actual_kappa, 0.025),
    upper_95 = quantile(actual_kappa, 0.975),
    .groups = "drop"
  ) %>%
  arrange(coherence, desc(trialval))

# 5. Render Styled Table
precision_summary_mix %>%
  mutate(across(where(is.numeric), ~ round(., 3))) %>%
  knitr::kable(
    format = "html",
    caption = "Response precision (kappa) by condition",
    col.names = c("Coherence", "Reward", "Median", "95% Lower", "95% Upper"),
    align = "llccc"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE
  ) %>%
  kableExtra::collapse_rows(columns = 1, valign = "top") %>%
  kableExtra::add_header_above(c(" " = 3, "95% Credible Interval" = 2))


```

As expected, there's a clear effect of coherence upon kappa (high-coherence trials produce a less dispersed distribution of responses); I won't do contrasts for this as it's pretty obvious.

Note also that with guesses filtered out, the model has found it even harder to estimate precision for the higher-coherence trials. This is another instance of identifiability issues when estimating *kappa* for tighter distributions (i.e., once a distribution is already 'spiked,' even massive increases in *kappa* result in almost no visible change to the response spread).

Within each level of coherence, is responding less precise for low-reward trials? In other words, are people making less of an effort when they aren't going to win big?

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing


kappa_draws_mix <- emmeans(fit_mixture_test, ~ trialval * coherence, dpar = "kappa1") %>%
  gather_emmeans_draws() %>%
  mutate(
    # mixture models in brms usually use a log link for kappa
    actual_kappa = exp(.value) 
  )

# 1. Re-extract and ensure we group by coherence before comparing reward levels
kappa_diffs_mix <- kappa_draws_mix %>%
  group_by(coherence) %>% # CRITICAL: Ensure coherence is the grouping variable
  compare_levels(actual_kappa, by = trialval, comparison = list(c("hi", "lo"))) %>%
  ungroup()

# 2. Re-run the plot (it should stay the same, but verify labels)
ggplot(kappa_diffs_mix, aes(x = actual_kappa, y = coherence, fill = coherence)) +
  stat_halfeye(.width = c(0.95, 0.66), point_interval = median_qi, alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  scale_fill_brewer(palette = "Oranges") +
  scale_y_discrete(limits = c("0.4", "0.2")) + 
  labs(title = "Change in precision across reward levels", x = "Difference in kappa (high - low reward)") +
  theme_minimal()

# 3. Re-run the summary table
kappa_diff_summary_mix <- kappa_diffs_mix %>%
  group_by(coherence) %>%
  summarize(
    median_diff = median(actual_kappa),
    lower_95 = quantile(actual_kappa, 0.025),
    upper_95 = quantile(actual_kappa, 0.975),
    prob_hi_better = mean(actual_kappa > 0),
    .groups = "drop" # Ensure groups are dropped for the kable
  )

# Check the table output now
kappa_diff_summary_mix %>%
  knitr::kable(digits = 3, format = "html") %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))
```

There's no effect when coherence is low, but we get an effect in the *wrong* direction when coherence is high (i.e., people are more precise for low-value trials than high-value ones). I'll explain more about this below, but I attribute this to an interaction with guessing proportion.

### Guessing proportion

Now, let's look the proportion of trials for which the model thinks people are guessing within each condition.

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing

cell_guess_mix <- emmeans(fit_mixture_test, ~ trialval * coherence, dpar = "theta2") %>%
  gather_emmeans_draws() %>%
  mutate(prob_guess = plogis(.value)) # Transform to 0-1 probability scale

# 2. Summary Table
zero_test_guess_summary <- cell_guess_mix %>%
  group_by(trialval, coherence) %>%
  summarize(
    median_prob = median(prob_guess),
    lower_95 = quantile(prob_guess, 0.025),
    upper_95 = quantile(prob_guess, 0.975),
    pd = as.numeric(bayestestR::p_direction(prob_guess)),
    .groups = "drop"
  )

# 3. Create the Kable Table
zero_test_guess_summary %>%
  mutate(across(where(is.numeric), ~ round(., 3))) %>%
  knitr::kable(
    caption = "Guessing probability by condition",
    col.names = c("Reward", "Coherence", "Median", "95% Lower", "95% Upper", "P-Direction"),
    align = "lccccc"
  ) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

# 4. Visualization
ggplot(zero_test_guess_summary, aes(x = trialval, y = median_prob, color = trialval)) +
  geom_pointrange(aes(ymin = lower_95, ymax = upper_95), size = 1, linewidth = 1.2) +
  facet_wrap(~ coherence, labeller = label_both) +
  scale_color_manual(values = c("hi" = "#ff6117", "lo" = "#009dff")) +
  labs(
    title = "Guessing probability by condition",
    x = "Trial Reward",
    y = "Guess probability (0-1)"
  ) +
  theme_minimal(base_size = 14)
```

The posterior estimates don't support a difference in guessing rate across coherence levels for high-value trials, but it *does* find that for low-reward trials, people guess more for high-coherence than low-coherence trials.

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing

# 1. Get raw cell probabilities (theta2)
cell_probs_guess <- emmeans(fit_mixture_test, ~ trialval * coherence, dpar = "theta2") %>%
  gather_emmeans_draws() %>%
  mutate(p = plogis(.value))

# 2. Calculate the difference: (0.2 Coherence - 0.4 Coherence) within each Reward
coherence_guess_effects <- cell_probs_guess %>%
  group_by(trialval) %>%
  compare_levels(p, by = coherence, comparison = list(c("0.2", "0.4"))) %>%
  ungroup() %>%
  mutate(reward_group = ifelse(trialval == "hi", "High Reward", "Low Reward"))

# 3. Summary Table
coherence_guess_summary <- coherence_guess_effects %>%
  group_by(reward_group) %>%
  summarize(
    median_diff = median(p),
    lower_95 = quantile(p, 0.025),
    upper_95 = quantile(p, 0.975),
    prob_increase = mean(p > 0), # Probability that harder task increased guessing
    .groups = "drop"
  )

# Render Table
coherence_guess_summary %>%
  mutate(across(where(is.numeric), ~ round(., 4))) %>%
  knitr::kable(
    format = "html",
    caption = "Effect of coherence on guess rate",
    col.names = c("Reward Level", "Median Diff", "Lower 95%", "Upper 95%", "P(Increase)"),
    align = "lcccc"
  ) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

# 4. Visualization
ggplot(coherence_guess_effects, aes(x = p, y = reward_group, fill = reward_group)) +
  stat_halfeye(
    point_interval = median_qi, 
    .width = c(0.95), 
    alpha = 0.8
  ) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Change in guess rate across coherence levels",
    x = "Change in guess rate (high - low coherence)",
    y = ""
  ) +
  theme_minimal(base_size = 14) +
  guides(fill = "none")

```

When coherence is low, the model doesn't find support for a difference in guessing rates based on trial reward. However, when coherence is high, it finds that people guess more often on low-value trials than high-value ones.

```{r}
#| echo: false      # Prevents the code itself from appearing in the document
#| message: false   # Prevents package loading messages and brms/Stan progress
#| warning: false   # Prevents warning messages
#| error: false     # Prevents error messages from printing

# 1. Get the Reward Effect (Hi - Lo) safely within each Coherence level
reward_guess_effects <- cell_guess_mix %>%
  group_by(coherence) %>%
  compare_levels(prob_guess, by = trialval, comparison = list(c("hi", "lo"))) %>%
  ungroup() %>%
  mutate(
    coherence_level = ifelse(grepl("0.2", coherence), "Low Coherence (0.2)", "High Coherence (0.4)"),
    coherence_level = factor(coherence_level, levels = c("High Coherence (0.4)", "Low Coherence (0.2)"))
  )

# 2. Summary Table
reward_guess_summary <- reward_guess_effects %>%
  group_by(coherence_level) %>%
  summarize(
    median_diff = median(prob_guess),
    lower_95 = quantile(prob_guess, 0.025),
    upper_95 = quantile(prob_guess, 0.975),
    prob_hi_lower = mean(prob_guess < 0), # Certainty that Hi Reward REDUCED guessing
    .groups = "drop"
  )

# Render Table
reward_guess_summary %>%
  mutate(across(where(is.numeric), ~ round(., 4))) %>%
  knitr::kable(
    format = "html",
    caption = "Effect of reward on guess rate",
    col.names = c("Coherence Level", "Median Diff", "Lower 95%", "Upper 95%", "P(Hi < Lo)"),
    align = "lcccc"
  ) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

# 3. Visualization
ggplot(reward_guess_effects, aes(x = prob_guess, y = coherence_level, fill = coherence_level)) +
  stat_halfeye(alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_fill_brewer(palette = "Greens") +
  labs(
    title = "Change in guess rate across reward levels",
    x = "Difference in guess rate (high - low reward)",
    y = ""
  ) +
  theme_minimal(base_size = 14) +
  guides(fill = "none")

```

## Takeaways

It's good to see that the model estimates the same pattern of value bias across conditions as the simpler (non-mixture) version.

However, there are some concerns here about estimation of kappa (precision) and theta (guess rate):

-   The model continues to struggle with estimating response precision when coherence is high, likely due to identifiability issues. When coherence is high, it also predicts that people are more precise for low-value trials than high-value ones, which is hard to justify empirically.

-   Guess rates look sensible for high-coherence trials, but they are identical regardless of trial reward when coherence is low. This is surprising and a bit concerning: as Christin and I have discussed, it likely reflects the model struggling to distinguish between guesses and low-precision responses to the target.

-   Furthermore, I suspect a relationship between some of the issues for theta and kappa: for instance, when considering high-coherence trials, maybe the model thinks that precision is higher in the low-value condition just because it has filtered out more trials as guesses.

## Next steps

While there are some things which have worked well in this mixture modelling, the data seem quite 'muddy' (i.e., there are a lot of datapoints which are quite hard to classify as genuine responses vs. guesses). This is partially because I added in our usually-excluded 'guessers' on this run, but it will be an issue we need to grapple with irrespective of that.

One really important question here: is this muddiness present *within* every individual distribution, or is it an effect that we see in the dataset as a whole because of *specific participants* who are contaminating the distributions (e.g., consistently guessing or using a strategy different from the instructed one)? As discussed with Christin, I think we will get better results if we rework our modelling approach to classify participants by strategy, then only estimate our parameters of interest for the subset of people who we're confident are doing the task as intended.
